Autoignition of Methyl Valerate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine

### 2017-06-29 methyl-valerate.tex
Updates

In general, there could be three likely sources of error in the models: missing
reaction pathways, incorrect values of the reaction rates, and incorrect values
for thermodynamic properties of the species. We have noted in
\cref{sec:reaction-mechanism-generator} that the RMG model has many more
reactions than the \cite{Dievart2013} model and the algorithm used in RMG
considers a substantial number of the possible pathways. This reduces the
possibility of missing reaction pathways affecting the model. Further detailed
studies are required to ensure that the RMG model includes all of the relevant
reaction pathways.

The second source of error may be incorrect reaction rate parameters, either
because the rates are specified incorrectly in the model (e.g., typos) or
because the rates are not well estimated by the typical analogy based-rules. It
should be noted that errors of this type may affect the model generated by
RMG---if the rates are not estimated correctly, reactions that are important in
reality may not be included in the model. Determining the accuracy of the
reaction rates used in the RMG and \cite{Dievart2013} models requires
further detailed studies of the models. Another related source of error could be
incorrect estimation of the pressure dependence of the reaction rates, which may
be particularly important for the isomerization reactions prevalent in
low-temperature chemistry.

The third source of error may lie in the estimation of the thermodynamic
properties of the species, particularly their heats of formation. We have begun
to analyze the possibility of this source of error by conducting a reaction
pathway analysis to determine which radicals are formed from the breakdown of
the fuel. The following analysis is conducted for a constant volume simulation
at \SI{700}{\K}, \SI{30}{\bar}, where the rates of production of the species
have been integrated until the time of \SI{20}{\percent} fuel consumption. The
results of this analysis are shown in \cref{fig:mv-structure} and
\cref{tab:mv-radicals} for the two models. The percentages shown in the
\cref{tab:mv-radicals} are the percent of the fuel destroyed to form a
particular fuel radical by all the reactions that can form that radical.

@@

\begin{figure}
    \begin{minipage}[t][][b]{0.3\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/Methyl_pentanoate.png}
        \caption{Structure of MV with carbon atoms labeled according to the convention used in \cref{tab:mv-radicals}}
        \label{fig:mv-structure}
    \end{minipage}\hfill%
    \begin{minipage}[t][][b]{0.68\textwidth}
        \centering
        \captionof{table}{Percent of MV destroyed to form fuel radical species with a hydrogen atom missing at the location indicated in the first column}
        \label{tab:mv-radicals}
        \begin{tabular}{cSS}
            \toprule
            Radical Site & {\cite{Dievart2013} [\si{\percent}]} & {RMG Model [\si{\percent}]} \\
            \midrule
            2 & 29.3 & 7.4 \\
            3 & 17.5 & 36.0 \\
            4 & 17.5 & 41.1 \\
            5 & 9.4 & 3.7 \\
            M & 26.3 & 11.8 \\
            \bottomrule
        \end{tabular}
    \end{minipage}
\end{figure}

>>>>

It is clear that neither model is able to predict the ignition delays of MV particularly well. In
addition to the poor agreement shown in \cref{fig:simulation-comparison}, the simulations for
\(P_C = \SI{15}{\bar}\), \(\phi=1.0\) and \(P_C = \SI{30}{\bar}\), \(\phi=\) \numlist{0.25;0.5;2.0}
showed substantial heat release during the compression stroke (i.e., the simulations are much too
reactive), and so these conditions aren't compared in \cref{fig:simulation-comparison}. We note
again that the model by \citet{Dievart2013} was validated for MV combustion only by comparison to
flame extinction limits, so the disagreement is not wholly surprising.

In general, there could be three likely sources of error in the models: missing reaction pathways,
incorrect values of the reaction rates, and incorrect values for thermodynamic properties of the
species. We have noted in \cref{sec:reaction-mechanism-generator} that the RMG model has many more
reactions than the \citet{Dievart2013} model and the algorithm used in RMG considers a substantial
number of the possible pathways. This reduces the possibility of missing reaction pathways affecting
the model. Further detailed studies are required to ensure that the RMG model includes all of the
relevant reaction pathways, which are outside the scope of this work.

The second source of error may be incorrect reaction rate parameters, either because the rates are
specified incorrectly in the model (e.g., typos) or because the rates are not well estimated by the
typical analogy based-rules. It should be noted that errors of this type may affect the model
generated by RMG---if the rates are not estimated correctly, reactions that are important in reality
may not be included in the model. Determining the accuracy of the reaction rates used in the RMG and
\citet{Dievart2013} models requires further detailed studies of the models, which are also outside
the scope of this work. Another, related, source of error could be incorrect estimation of the
pressure dependence of the reaction rates, which may be particularly important for the isomerization
reactions prevalent in low-temperature chemistry.

The third source of error may lie in the estimation of the thermodynamic properties of the species,
particularly the fuel radicals. By conducting a reaction pathway analysis to determine which
radicals are formed from the breakdown of the fuel, we can analyze the proportion of each radical
formed as the fuel breaks down during the autoignition process. The following analysis is conducted
for a constant volume simulation at \SI{700}{\K}, \SI{30}{\bar}, and the stoichiometric equivalence
ratio, where the rates of production of the species have been integrated until the time of
\SI{20}{\percent} fuel consumption. The results of this analysis are shown in
\cref{fig:mv-structure} and \cref{tab:mv-radicals} for the two models. The percentages shown in the
\cref{tab:mv-radicals} are the percent of the fuel destroyed to form a particular fuel radical by
all the reactions that can form that radical.

@@

\begin{center}
    \captionsetup{type=figure}
    \includegraphics[width=90mm]{figures/Methyl_pentanoate.png}
    \caption{Structure of MV with carbon atoms labeled according to the convention used in \cref{tab:mv-radicals}}
    \label{fig:mv-structure}
\end{center}

\begin{center}
    \captionsetup{type=table}
    \caption{Percent of MV destroyed to form fuel radical species with a hydrogen atom missing at
    the location indicated in the first column and \cref{fig:mv-structure}}
    \label{tab:mv-radicals}
    \begin{tabular}{cSSSS}
        \toprule
        Radical Site & {\citet{Dievart2013} [\si{\percent}]} & {RMG Model [\si{\percent}]} & & \\
        \midrule
        2 & 29.3 & 12.5 & 27.7 & 15.6\\
        3 & 17.5 & 12.2 & 17.8 & 20.1\\
        4 & 17.5 & 50.6 & 17.8 & 23.2 \\
        5 & 9.4 & 3.9 & 10.1 & 2.7 \\
        M & 26.3 & 20.8 & 26.6 & 38.4\\
        \bottomrule
    \end{tabular}
\end{center}
